---
title: "Project-2 Data Wrangling"
author: "Ben Baumer"
format: html
---

# Learning Goals

- To further sharpen data wrangling and data visualization skills
- To learn how to collaborate using modern workflows

# Technical Skills

- `dplyr`
- `tidyr`
- GitHub


## Project

You will work in a group of two or three (groups posted to Slack and Moodle) to analyze [Federal Election Commission](http://www.fec.gov) data contained in [the `fec20` package](https://www.github.com/baumer-lab/fec20), and report your findings in a short write-up. The topic is up to you and your group. The best projects will:

-   discuss an interesting and well-motivated topic
-   involve some non-trivial data wrangling (e.g. not just a bunch of `mutate()`s)
-   provide a well thought-out, informative analysis
-   convey some sort of insight
-   be well-written
-   consist of 400--600 words

# GitHub

You will use GitHub to collaborate with your partner(s) on this assignment. 


# Submission

1. Push all commits to GitHub
2. The final commit should have a rendered `.html` file with a message teeling me the project is complete. 

# Data

You will have to install the `fec20` package.

```{r, echo=TRUE, eval=FALSE, include=FALSE}
install.packages("remotes")
remotes::install_github("baumer-lab/fec20")
```

These data come from the [Federal Election Commission](https://www.fec.gov/data/browse-data/?tab=bulk-data), and are based on the 2019--2020 federal election cycle. These data were collected using the [`fec20`](https://github.com/baumer-lab/fec20) package for R. 

There are several tables present. To load them, switch to the `mp2-group-x` project in RStudio, pull, and run the following:

```{r, echo=TRUE, eval=FALSE}
library(fec20)
```

> **Be sure to read the [supporting documentation for these data](https://www.fec.gov/data/browse-data/?tab=bulk-data)!!** It is your responsibility to know what you are looking at!



