---
title: "map() practice"
author: "Schwab"
editor: visual
---

```{r}
library(tidyverse)
library(mosaicData)
```

Doing problems 1 and 3 from [here](https://mdsr-book.github.io/mdsr3e/07-iteration.html#iteration-exercises)

**Problem 1 (Easy)**: Use the `HELPrct` data from the `mosaicData` to calculate the mean of all numeric variables (be sure to exclude missing values).\

```{r}
HELPrct |>
  summarise(
    across(.cols = where(is.numeric),
           .fns = ~mean(. , na.rm = TRUE))
  )
```

**Problem 3 (Medium)**: Use routines from the `purrr` package and the `HELPrct` data frame from the `mosaicData` package to fit a regression model predicting `cesd` as a function of `age` separately for each of the levels of the `substance` variable. Generate a formatted table (with suitable caption) of results (estimates and confidence intervals) for the slope parameter for each level of the grouping variable.

(Hint: Use `group_by()` and `group_modify()` to fit the regression model on each part. Note that `broom::tidy()` is useful in having the output of the model be a data frame.)

```{r}
HELPrct |> select(substance, cesd, age) |>
group_by(substance)|>
  group_modify(.f = ~broom::tidy( lm(cesd~age, data = .))) |> 
  filter(term =="age") |>
  mutate(CI_low = estimate - 2*std.error, CI_high = estimate + 2*std.error) |>
  select(-statistic, - p.value, - std.error) |>
  kableExtra::kable(caption = "The likelyhood someone uses a substance increates with age.", digits = 1)
```

Activity 2

Make a function that takes a df and a variable then produces a scatter plot of that variable vs cesd. Use it to plot multiple graphs of the first 6 numeric variables HELPrct. [Hint: check out the extended example from 7.7](https://mdsr-book.github.io/mdsr3e/07-iteration.html#extended-example-factors-associated-with-bmi)

```{r}
library(tidyverse)
library(mosaicData)

my_func <- function(data,variable2){
  data |>
    ggplot(aes(x = cesd, y = .data[[variable2]]))+
    geom_point()
}

my_func(HELPrct, "age")

numeric_names_only <- HELPrct |>
  select("age":"d1")|>
  names()


map(.x = numeric_names_only,
    .f = my_func,
    data= HELPrct
      )|>
  patchwork::wrap_plots(ncol = 2)
```

```{r}
library(tidyverse)
library(mosaicData)

my_func <- function(data,variable2){
  data |>
    ggplot(aes(x = cesd, y = variable2 ))+
    geom_point()
}



```

```{r}
library(NHANES)

bmi_plot <- function(.data, x_var) {
  ggplot(.data, aes(y = BMI)) +
    aes_string(x = x_var) + 
    geom_jitter(alpha = 0.3) + 
    geom_smooth() + 
    labs(
      title = paste("BMI by", x_var),
      subtitle = "NHANES",
      caption = "US National Center for Health Statistics (NCHS)"
    )
}
bmi_plot(NHANES, "Age")

c("Age", "HHIncomeMid", "PhysActiveDays", 
  "TVHrsDay", "AlcoholDay", "Pulse") |>
  map(bmi_plot, .data = NHANES) |>
  patchwork::wrap_plots(ncol = 2)
```
