---
title: "File Paths and Intro to Maps"
author: "Schwab"
editor: visual
---

Overview of the course

-   Introduction to Programming Language for Data Science (R) and Computing Environment (RStudio)

    [Short History of R](https://nics-github.github.io/SDS192/course-materials/lectures/Brief-R-history.html)

-   Data Science Life Cycle

    [DS life cycle](https://towardsdatascience.com/data-science-101-life-cycle-of-a-data-science-project-86cbc4a2f7f0)

-   Grammar of Graphics

    [Hadley Wickham](https://byrneslab.net/classes/biol607/readings/wickham_layered-grammar.pdf)

-   Data Collection

[Collect student data here](https://forms.gle/FQhAau8GuKURoH2p9)

-   Web Scraping and Importing Data

```{r import_student_data}
library(googlesheets4)
library(janitor)

student_data <- read_sheet('https://docs.google.com/spreadsheets/d/1_QCcgjkGeGpBLtn5RrmM1SLhYq7xkgEhkMsazzZlJIs/edit?resourcekey=&gid=347243538#gid=347243538')

student_data <- clean_names(student_data)

```

-   Data Visualization

```{r}
# make simple plot of student data. 
library(tidyverse)

student_data |>
  ggplot(mapping = aes(x = what_is_your_height_in_inches) )+
  geom_histogram()+
  labs(
    title = "Students' height in inches"
  )
```

-   Data Wrangling

```{r wrangle_s_d}
# wrangle student data. edit plot
student_data |>
  select(briefly_describe_any_previous_experience_you_have_with_r) |>
  rename(r_experiance = briefly_describe_any_previous_experience_you_have_with_r )|>
  group_by(r_experiance)|>
  summarize(
    sum()/n()
  )
  
```

-   Data Management

Keeping yourself organized!

-   Data Ethics

-   Iteration

```{r}
# have r do many repeate steps for us. 
```

-   Introduction to Database Querying with SQL

Do all the same stuff we learned above with SQL.

[Simple Query?](https://www.w3schools.com/sql/trysql.asp?filename=trysql_editor)

-   Further Topics based on time and student interest may include:

    -   Dynamic and Customized Data Graphics

    -   Geospatial Data and Maps

    -   Learning with AI
