---
title: "Activity - ggplot 2"
subtitle: "Meet ggplot"
output: 
  tufte::tufte_html:
    css: ../hw.css
    tufte_variant: "envisioned"
    highlight: pygments
link-citations: yes
---

```{r include = FALSE}
knitr::opts_chunk$set(
  eval = FALSE,
  out.width = "80%",
  fig.asp = 0.618,
  fig.width = 10
)
library(tidyverse)
library(openintro)
library(ggrepel)
```

We'll use the [mammls](https://www.openintro.org/data/index.php?data=mammals) dataset.

Here's a quick rundown of how things work in ggplot. We'll learn more as we go. 

Make a chunk just to look at the data. 
```{r}
mammals
```
Make another one to dotplot the data. You need:
- a pipe |>
- ggplot() with some aesthetics
- a **+**, 
- geom_point()
```{r}
mammals |>
  ggplot(aes(x=total_sleep, y =brain_wt)) +
  geom_point()
  
```
Lets add a color aesthetic.

```{r}
mammals |>
  ggplot(aes(x=total_sleep, y =brain_wt, color = exposure)) +
  geom_point()
  
```

Let's add a facet()

```{r}

mammals |>
  ggplot(aes(x=total_sleep, y =brain_wt, color = exposure)) +
  geom_point() +
  facet_wrap(~danger)
  
```

How do you know which aesthetics can be added? 

**?geom_point()**

How do you choose a facet? 

Choose something witih finite options. 

Also see the ggplot cheatsheet under [Resources](./Resources.html)

