---
title: "Color exercise"
author: "Schwab"
format: html
editor: 
  markdown: 
    wrap: 72
---

Let's play around with some colors.

```{r message=FALSE}
library(openintro)
library(tidyverse)
library(RColorBrewer)
library(ggthemes)
```

Here's the [presidents data
set](https://www.openintro.org/data/index.php?data=president). Let's
play with themes. Add theme_bw(), theme_economist(), theme_dark(),
theme_classic()

```{r presidents}
president |>
  ggplot(aes(x=start, y =end, color = party) )+
  geom_point()+
  theme_bw()

```

```{r}
display.brewer.all()
```

That's alittle hard to see in the .rmd. [Here are the color brewer
choices](https://r-graph-gallery.com/38-rcolorbrewers-palettes.html)

Add scale_fill_brewer

```{r x-party-fill-party}
president |>
  ggplot(aes( x = party,fill=party) )+
  geom_bar()+
  scale_fill_brewer(palette = "Purples") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
           
```

Does color add anything to the graph above?

Let's make a count plot of [assortative
mating](https://www.openintro.org/data/index.php?data=assortive_mating).
[Notice we are adding a shape
aesthetic.](https://ggplot2.tidyverse.org/articles/ggplot2-specs.html)

```{r assortative-mating}
assortative_mating |>
  ggplot(aes(x = self_male, y= partner_female, fill=self_male),)+
  geom_count(aes(color = partner_female), shape=23,stroke=1.5)  + 
  scale_fill_manual(values = c("blue","brown","green")) +
  scale_color_manual(values =  c("blue","brown","green")) +
  theme_bw()
```

How do you feel about color in the above graph?

Try this:

-   Make sure the above libraries are loaded.

-   Take the [cpr data
    set](https://www.openintro.org/data/index.php?data=cpr) from the
    OpenIntro package

-   Make a geom_bar() graph for the variable `group`

-   Use the fill aesthetic to color the bars by outcome.

-   Add a scale_fill_brewer(palette="" ) to play with different [brewer
    choices](https://r-graph-gallery.com/38-rcolorbrewers-palettes.html).

```{r make-a-cpr-plot}
cpr 
```

Copy most of the chunk you made above.

-   Add scale_fill_manual(values=c("black","white")) to manually set two
    colors for outcomes. Change black and white to other colors. [Here's
    a list of color choices in
    R.](https://derekogle.com/NCGraphing/resources/colors)

-   Copy most of the chunk again and try scale_fill_fivethirtyeight()
