---
title: Pivot Practice
author: Nic Schwab 
---

```{r setup,  message=FALSE}
knitr::opts_chunk$set(message = FALSE,echo=FALSE)

library(fivethirtyeight)
library(tidyverse)
#When you get to problem 3 uncomment this library.
#library(countrycode) 
options(scipen = 10 )

```

We'll tidy up the `drinks` data frame from `fivethirtyeight` and the `billboard` dataframe from `tidyr`. Then we'll join the codelist data frame to the world_bank_pop to show how the world population has changed over time.

## Problem 1: Heavy Drinkers

Goal: Pivot longer the `drinks` dataframe. Make a bar plot of the top four most intoxicated countries (I ended up using geom_col). 

-   `rename()` the columns to drop the "*servings".*

-   Drop the `total_litres_column`.

-   Pivot and save the new data frame. 

```{r pivot_drinks}

  
```

-   Plot a bar graph for the top 4 alcohol drinking countries and set position = "dodge" in the atheistic.

```{r graph_top_4}

```


## Problem 2: Toping the charts. 

Goal: Find out which artist topped the charts longest in the year 2000. 

Take the `billboard` data frame and tell me which song was in the one spot of the charts for longest. 

Hint: If you tell R cols= wk1:wk76 it'll know you mean all columns between those weeks.

```{r topping_the_charts}


```



## Problem 3: Populations are growing. 

Goal: We'll use the `codelist` dataframe and `world_bank_pop` dataframe to make a point plot of each continents populations by year. You will need to install the country code  run install.packages("countrycode") in the console (not in the markdown file) to install the `countrycode` package.


Using both an inner_join(), a pivot_longer(), and a few verbs make a point plot showing how the total population of each continent changes between 2000 and 2017.

```{r codelist}
# This will limit the codelist dataframe to just three easy to read rows. 

codelist_subset <- codelist |> 
  select(continent,country.name.en, cowc)
```

```{r population_growth_by_continents}
#This will get you started with the point plot. 


```
