---
title: "Stop Question Frisk "
subtitle: "An Introduction to Wrangling"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
---

## Stop, Question, Frisk

-   Started in the 1990s

-   New York City Police Practice

    -   Police need "[reasonable suspicion](https://supreme.justia.com/cases/federal/us/392/1/)."

-   Leads to mistrust in the police by these groups.

## More about about me

I'm not Black.

I am Hispanic.

I look White.

I am not from New York City.

In this data analysis I will have limited perspective.

## Race Exists

This is an acknowledgement that the construct of race exists and plays a part in the way human being interact with each other.

## Schedule

::: incremental
1.  Introduce you to the data.

2.  Pose a question.

3.  Give you [data frames]{.underline} with data to answer the question one way.

    i.  You'll order them with a partner.

    ii. You'll make up [pseudocode]{.underline} to explain data transformations.

    iii. I'll give you the real code.

4.  We'll repeat the process of for 2021 data in R.
:::

# Questions so far?

## The Data

[Here is](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page) NYC's stop and frisk data.

Let's consider 2011 data

::: nonincremental
-   Each column is a variable

-   and each row is a stop
:::

*The variable names are inconsistent across years.*

## SQF History

```{r stop-frisk-years}
#| echo: TRUE
#| warning: FALSE
#| include: FALSE

library(tidyverse)
library(readxl)
options(scipen = 10)



```

```{r }
#| eval: False

# Read in each csv file. This is a lot 

# With purr csv

file_paths_csv <- fs::dir_ls("../data/2003_2015_csv_files")

file_contents_csv<- file_paths_csv |>
  map(function (path){
    read_csv(path)
  })

# with purr excel

file_paths_xl <- fs::dir_ls("../data/2016_2022_xl_files")

file_contents_xl<- file_paths_xl |>
  map(function (path){
    read_excel(path)
  })

# https://www.youtube.com/watch?v=An1bUIg-nVM

#See the number of rows i.e stops in each year
pre_2016 <- sapply(file_contents_csv, nrow)
post_2016 <- sapply(file_contents_xl, nrow)


```

```{r graph-of-stop}

years <- c(2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013, 2014, 2015, 2016, 2017, 2018, 2019, 2020, 2021, 2022 )
stops_frisks <- c(160851, 313523, 398191, 506491, 472096, 540302, 581168,  601285, 685724, 532911,  191851, 45787, 22563, 12405, 11629,  11008,   13459,9544 , 8947,  15102)  

sqf <- data.frame(years,stops_frisks)



sqf |> 
  ggplot(aes(years,stops_frisks)) +
  
  #geom_rect(aes(xmin=2002,xmax=2014,ymin=0,ymax=Inf),fill="#003585",alpha=0.01)+
  #geom_rect(aes(xmin=2014,xmax=2021,ymin=0,ymax=Inf),fill="#FFFFFF",alpha=0.01)+
  #geom_rect(aes(xmin=2021,xmax=2022,ymin=0,ymax=Inf),fill="#FF6600",alpha=0.01)+
  geom_line()+
  xlab("Year")+
  ylab("Citizens Stopped ('000s)")+
  labs(title="")+
  theme_minimal()+
  scale_y_continuous(breaks = c(0,350000,700000), labels = c("0","350","700"))+
  scale_x_continuous(breaks = c(2003,2013,2022))+
  geom_text(label="Bloomberg", x = 2004.5, y = 40000, color="gray35" ) +
  geom_text(label="de Blasio", x = 2016, y = 40000, color="gray35" ) +
  geom_text(label="Adams", x = 2022, y = 40000, color="gray35" ) +
  geom_text(label="NYCLU \n suit", x = 2013, y = 650000, color="gray35" ) +
  geom_vline(xintercept=2012, linetype = 3 ) +
  theme(
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    axis.line = element_line(color = "black"),
    axis.text=element_text(size=12)
  
  )
```

## Questions

::: incremental
  1.  Which racial groups were most likely to be stopped by police officers in 2011?

  2.  What percent of the stops resulted in an arrest by race in 2011?

  3.  How do the answers to these questions change in 2021?

  4.  You will likely have more questions as we analyze the data.
  
:::

## Your data frames

1.  Data envelopes for **2011 Data**

-   Put the data frames in order.

-   Be sure you can explain the process from one df to another. "psuedocode"

-   Questions about the data?

2.  Arrow envelopes **functions**

-   After the frames are in order put the code arrows between each frame.

-   Questions about the code?

## Conclusions from 2021 data

Here are our questions again 

::: incremental
  1.  Which racial groups were most likely to be stopped by police officers in 2011?

  2.  What percent of the stops resulted in an arrest by race in 2011?

  3.  How do the answers to these questions change in 2021?

  4.  You will likely have more questions as we analyze the data.
  
:::

## For comparison:

Here's the demographics of NYC[^1]:

[^1]: https://www.census.gov/quickfacts/fact/table/newyorkcitynewyork/PST045222

-   Black: 23.4 %

-   Latino: 28.9%

-   White: 39.8%

## Summarize functions:

New functions:

-   select(), mutate(), arrange(), rename()

Old functions:

-   names(), recode(), count()

## Let's compare to 2011.

Make a new r project.

Download 2011 csv from [this site.](https://www.nyc.gov/site/nypd/stats/reports-analysis/stopfrisk.page) 

Make a new document and put this in a chunk.

```{r}
#| eval: false
#| echo: true
#| code-fold: true

# This code will read in the 2011 SQF data from the internet

# makes a temp file
temp <- tempfile()

# downloads the file from the url and puts it in the temp location
download.file("https://www.nyc.gov/assets/nypd/downloads/zip/analysis_and_planning/stop-question-frisk/sqf-2011-csv.zip",temp)

# This makes a connection to the temp filea and reads the csv in the temp location 
data <- read.csv(unz(temp, "2011.csv"))

# This breaks the connection to the temp file (important if you are reading other files later. )
unlink(temp)
```

## Answer the questions.

Figure out the variable names and do a similar analysis.

> Are Black, and Hispanic groups more likely to be stopped by police officers than other groups in 2011? 

> Does this change as the total counts of stops decreases to 2021 levels?

## Two more verbs

-   filter() - subsets the data based on requirements.

-   mutate() - adds a new column.
