---
title: "Course Overview"
subtitle: "SDS 192"
author: "Schwab"
editor: visual
format: 
  revealjs: 
    theme: beige
execute: 
  echo: true
---

## Introduction to Programming Language for Data Science (R) and Computing Environment (RStudio)

-   Take a look at R Studio together

-   [Short History of R](https://nics-github.github.io/SDS192/course-materials/lectures/Brief-R-history.html)

## Data Science Life Cycle

[DS life cycle](https://towardsdatascience.com/data-science-101-life-cycle-of-a-data-science-project-86cbc4a2f7f0)

## Grammar of Graphics

[Hadley Wickham](https://byrneslab.net/classes/biol607/readings/wickham_layered-grammar.pdf)

## Data Collection

[Collect student data here](https://forms.gle/FQhAau8GuKURoH2p9)

## Web Scraping and Importing Data ( and a tiny bit of cleaning)

```{r import_student_data}
library(googlesheets4)
library(janitor)

student_data <- read_sheet('https://docs.google.com/spreadsheets/d/1k9qA7XldtHB8-FZjmEymoUoeZJArS4tGy5gyVsvyLB0/edit?usp=sharing')

student_data <- clean_names(student_data)

```

## Data Visualization

```{r}

# make simple plot of student data. 
library(tidyverse)

student_data |>
  ggplot(mapping = aes(x = what_is_your_height_in_inches) )+
  geom_histogram()+
  labs(
    title = "Students' height in inches"
  )
```

## Data Wrangling

```{r wrangle_s_d}
# wrangle student data. edit plot
student_data |>
  
  select(briefly_describe_any_previous_experience_you_have_with_r) |>
  
  rename(r_experiance = briefly_describe_any_previous_experience_you_have_with_r )|>
  
  group_by(r_experiance)|>
  
  summarize(
    
    sum()/n()
    
  )
  
```

## Data Management

```         
-   Security
-   Keeping yourself organized!
-   How Data is housed
```

## Data Ethics

How we use data without doing harm.

## Iteration

```{r}
# have r do many repeat steps for us. 

print(student_data$one_word_summer_highlight)

```

## Verson Control

With git and github.

## Introduction to Database Querying with SQL

Wrangle some data with the classic SQL.

[Simple Query?](https://www.w3schools.com/sql/trysql.asp?filename=trysql_editor)

## Further Topics based on time and student interest may include:

```         
-   Dynamic and Customized Data Graphics

-   Geospatial Data and Maps

-   Learning with AI
```
