---
title: "Static Maps w/ ggplot"
author: "Schwab"
format: 
  revealjs: 
    theme: beige
editor: visual
execute: 
  echo: true
---

## Libraries

```{r setup}
library(tidyverse)
library(sf)
library(tidygeocoder)
library(maps)

```

## tidygeocoder

Let's find the latitude and longitude for a couple of places.

```{r}

addresses<- tibble(address=c("Smith College", "2 Tyler Ct, Northampton, MA 01060", "Holyoke Community College")) |>
geocode(address, method="osm") |>
  mutate(name = c("Smith College","McConnell Hall","Holyoke Community College"))

addresses
```

## My first map

Let's plot those with ggplot.

```{r}
ggplot(addresses) +
  geom_point(aes(long,lat)) 
  
  

```

Not very exciting. We'll add it it in a moment.

## Mass Map data

Let's build a map of Massachusetts with the map_data() function

```{r}
# map_data comes from ggplot2

ma_counties <- map_data("county", "massachusetts") |>
  select( long, lat, group, id = subregion)

head(ma_counties)

```

## Mass Map

Notice there are a lot of points for barnstable. Those are the vertices of a polygon. Let's draw the polygon with the points.

```{r}
ma_counties |> ggplot(aes(long,lat,group=group))+
  geom_polygon(fill = "white", colour = "grey50")+
  geom_point()  
  #coord_sf()
```

## Mapping projections

Coordinate Reference Systems (CRS) - Is the type of map projection employed To get a 3D Earth to a 2D surface.

Which one will coord_sf default to?

```{r}
?coord_sf
```

[This is the crs commonly used in gps.](https://epsg.io/4326)

## West wing explanation

[West Wing](https://www.youtube.com/watch?v=vVX-PrBRtTY)

## Adding the 4326 crs

```{r}
ma_counties |> ggplot(aes(long,lat,group=group))+
  geom_polygon(fill = "white", colour = "grey50")+
  geom_point()+
  coord_sf()
```

## Add some cities

Let's grab just Massachusetts cities from the data frame us.cities.

```{r}
# us.cities comes from the maps package

ma_cities <- us.cities |>
  filter(country.etc =="MA") |>
  select(name, long, lat)

head(ma_cities)
```

## Exercise:

Plot MA cities onto our county map. Then clean up our map.

```{r}
ma_counties |> ggplot(aes(long,lat))+
  geom_polygon(fill = "white", colour = "grey50",aes(group=group)) +
  #geom_point(data=ma_cities,.......)
  
  #This is the default mapping. Think GPS
 coord_sf(default_crs= st_crs(4326) )
          
          
          # This is Mercator
          # st_crs(3857))
```

## Shape files

Often times map data is saved as a shape file.

The US government has shape files [here](https://www.census.gov/geographies/mapping-files/time-series/geo/carto-boundary-file.html).
