---
title: "Anonymous Functions"
author: "Schwab"
editor: visual
format: 
  revealjs:
    theme: beige
execute: 
  echo: true
---

## Digression

Anonymous functions are functions without a name.

## A named function

```{r}
library(stringr)

#This function has a name, upper_case

upper_case <-  function(word){
  str_to_upper(word)
}

upper_case("hi")
```

## Anonymous functions

This is the same function without a name

```{r}
function(word) str_to_upper(word)
```

Its harder to use by itself, we often use them in combination with other functions.

## Using the anonymous function

The function after .f will only be used once, we don't need to name it.

```{r}
words <- c("I", "Like", "to", "eat", "cheese", "!")
purrr::map_chr(
  .x = words,
  .f = function(word) str_to_upper(word)
)
```

## Name an anonymous function

You can name an anonymous function.

```{r}
# This works with map_chhr() because str_to_upper is vectorized

upper_case_anon <- function(word) str_to_upper(word)
upper_case_anon(words)
```

## Does this work?

```{r}
# This works because str_to_upper is vectorized

upper_case_anon <- function(word) 
  str_to_upper(word)

upper_case_anon(words)
```

Yes, but it can get confusing.

## This does not work

```{r, eval=FALSE}
library(tidyverse)

confusing_function <- function(word) 
  upper_case_words<- 
    str_to_upper(word)
## This should lower the word's case, but it doesn't because its not part of the function.  
  str_to_lower(upper_case_words)

confusing_function(words)
```

## Last example fixed with

```{r}
confusing_function <- function(word){ 
  upper_case_words<- 
    str_to_upper(word)
  ## Now it is part of the function.  
  str_to_lower(upper_case_words)
}
  
confusing_function(words)
```

## Summary

Use an anonymous function to do one thing.

If it get complicated give some {}

## More anonymous?

Since R version 4.1, May 2021.

Just use a `\(argument) operations`

```{r, eval=FALSE}
# This is also an anonymous function. 
\(word) str_to_upper(word)

map_chr(.x = words,
    .f = \(word) str_to_upper(word))
```

## Deep dive

There are actually lots of ways of making anonymous functions.

[Here are more.](https://stackoverflow.com/questions/52492462/define-an-anonymous-function-without-using-the-function-keyword) I like the idea of the lambda package.

## SDS 270 for more on functions in R
