---
title: "Notes on Style"
author: "Schwab"
format: 
  revealjs:
    theme: beige
    controls: 
      true
execute: 
  echo: true
editor: visual
---

## The Tidyverse Style Guide

[Tidyverse style guide](https://style.tidyverse.org/index.html)

## 1. files 

I will use underscores `_`, not spaces to name your files. You can also use `-`

-   Avoid special symbols.

-   Called [snake case.](https://en.wikipedia.org/wiki/Snake_case)

### good name

`my_file.qmd`

### poor name

`my file!.qmd`

## 2. Syntax 

syntax is how the code looks

-   variables should be in snake case and lower case

-   variables should be nouns

-   functions should be verbs

-   names should be meaningful and concise.

-   Don't reuse built in R names (True, False, c, n)

```{r}
# good
first_name <- "Nic"

# not so good
LASTNAME <- "Schwab"
```

## 2.5 More Syntax

Consider spacing and commas like in english writing.

More space for reability is ok. Don't over do it.

```{r}
# good
x <- 1
y <- 2
total <- sum(c(x, y))
  
# Not as good
new_total<-x+y-total 

# Maybe ok
sum( c( x , y , total ) )

```

## Functions

Indent long lines (R does this automatically)

We don't need return, R automatically returns the last line.

```{r}
add <- function(a = "first number",
                b = "second number",
                c = "third number") {
  a + b + c
  # As usual code is indented by two spaces.
}
```

## Pipes and plus

In general `|>` and `+` should end lines.

We will discuss these operators more later.
