---
title: "Frequency Plots"
author: "Schwab"
format: 
  revealjs:
    theme: beige
editor: visual
execute: 
  echo: true
---

## Libraries

```{r}
library(tidyverse)
```

## Recall Variable Types

Categorical Variables in R are called:

-   Character variables

-   Factors

Numeric Variables in R are called:

-   Integer

-   Double

## Create Histograms

To see the spread, center and shape of a numeric variable.

```{r}
starwars |>
  ggplot(aes(x= height )) +
  geom_histogram()
```

## Histograms

1.  Make bins for the numbers
2.  Counts the occurrence of a value within bins
3.  The height is the number of counts.

We can adjust the bin or bin size.

## Create Bar graph

Visualize the counts of a categorical variable

```{r}
starwars |>
  ggplot(aes(x = hair_color))+
  geom_bar()
```

## Two bar graphs

Use geom_bar() is data is not precounted

Use geom_col() if data is precounted.

```{r}
starwars |>
  count(hair_color, name = "no_of_characters") |>
  ggplot(aes(x= hair_color, y= no_of_characters)) +
  geom_col()
```

## Create Facets

We can split graphs by another categorical or discrete variable.

```{r}
starwars |>
  ggplot(aes(x = hair_color))+
  geom_bar() +
  facet_wrap(~gender)+
  theme(axis.text.x = element_text(angle = 90)) # rotating text so its easier to read
```

## Too many facets

Be careful too many facets can obscure the visualization.

```{r}
starwars |>
  ggplot(aes(x = hair_color))+
  geom_bar() +
  facet_wrap(~species) +
  theme(axis.text.x = element_text(angle = 90))
```

## Try one

Facet starwar's characters' height by their gender.
