---
title: "maps-ggplot"
author: "Schwab"
format: html
editor: visual
---

```{r setup}
library(tidyverse)
library(mdsr)
library(sf)
library(ggspatial)
library(leaflet)
library(tidygeocoder)
library(mapproj)
library(maps)

CholeraDeaths_Book <-mdsr::CholeraDeaths
st_crs(CholeraDeaths_Book) = st_crs(CholeraDeaths_Book) 
```

Let's find the latitude and longitude for a couple of places.

```{r}

addresses<- tibble(address=c("50 Payson Ave, Easthampton, MA","Holyoke Community College, Holyoke,MA")) |>
geocode(address, method="osm") |>
  mutate(name = c("City Hall","HCC"))

addresses
```

Let's plot those with ggplot.

```{r}
ggplot(addresses) +
  geom_point(aes(long,lat)) 
  #coord_quickmap()
  

```

Not very exciting. We'll add it it in a moment.

Let's build a map of Massachusetts with the map_data() function

```{r}
ma_counties <- map_data("county", "massachusetts") %>% 
  select( long, lat, group, id = subregion)

head(ma_counties)

```

Notice there are a lot of points for barnstable. Those are the vertices of a polygon. Let's draw the polygon with the points.

```{r}
ma_counties |> ggplot(aes(long,lat,group=group))+
  geom_polygon(fill = "white", colour = "grey50")+
  geom_point()+
  coord_quickmap()
```

Let's grab just Massachusetts cities from the data frame us.cities.

```{r}
ma_cities <- us.cities %>% 
  filter(country.etc =="MA") |>
  select(name, long, lat)

head(ma_cities)
```

Exercise: Plot MA cities onto our county map.

```{r}
ma_counties |> ggplot(aes(long,lat))+
  geom_polygon(fill = "white", colour = "grey50",aes(group=group)) +
  #geom_point(data=ma_cities,.......)
  
  coord_quickmap()
```

```{r}
class(ma_counties)
class(ma_cities)
```

Let's get some more accurate government data.

[Go here](https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html) and download some shape files.

Saving them will be a bit of work. First we need to find those files.

```{r}
states <- read_sf("~/Documents/MTH_190_Intro_to_DS/course-materials/cb_2021_us_state_500k") |>
  sf::st_transform('+proj=longlat +datum=WGS84')

counties <- read_sf("~/Documents/MTH_190_Intro_to_DS/course-materials/cb_2021_us_county_5m") |>
  st_transform('+proj=longlat +datum=WGS84')
```

```{r}

leaflet(states) |>
  addTiles() |>
  addPolygons()
```

```{r}
Just_MA <-states |> 
  filter(STUSPS =="MA") 
  
leaflet(Just_MA) |>
  addTiles() |>
  addPolygons()
```

```{r}
Just_MA_counties <- counties |> filter(STUSPS == "MA")

leaflet() |>
  addTiles() |>
  addPolygons(data = Just_MA_counties) |>
  addMarkers(data = addresses, lng=~long, lat=~lat, popup = ~name)
  
```

```{r}
Just_MA_counties <- counties |> filter(STUSPS == "MA")

leaflet() |>
  addTiles() |>
  addPolygons(data = Just_MA_counties) |>
  addMarkers(data = addresses, lng=~long, lat=~lat, popup = ~name, ) |>
  addCircleMarkers(data = ma_cities, lng=~long, lat=~lat, popup = ~name, color='black',clusterOptions = markerClusterOptions() ) 
  
```

```{r}
leaflet() |>
  addTiles() |>
  addPolygons(data = Just_MA_counties) |>
  addMarkers(data = addresses, lng=~long, lat=~lat, popup = ~name, ) |>
  addCircleMarkers(data = ma_cities, lng=~long, lat=~lat, popup = ~name, color='black',clusterOptions = markerClusterOptions() ) 
```

Download some more shape files from [here](https://www.census.gov/geographies/mapping-files/time-series/geo/cartographic-boundary.html).

Exercise: Plot the Places we should care about from MA on our map.

Exercise: Plot the American Indian Tribal Subdivisions.
