---
title: "maps-leaflet-2"
author: "Schwab"
format: revealjs
editor: visual
---

```{r}
dsn <- fs::path("~","Documents","MTH_190_Intro_to_DS","course-materials", "SnowGIS_SHP")
list.files(dsn)
```

```{r}
st_layers(dsn
          )
```

```{r}
CholeraDeaths <- st_read(dsn, layer = "Cholera_Deaths")
```

```{r}
class(CholeraDeaths)
```

```{r}
CholeraDeaths
```

```{r}
CholeraDeaths_Book<- mdsr::CholeraDeaths
```

```{r}

ggplot((CholeraDeaths_Book)) +
  geom_sf(  )
```

```{r}
ggplot(CholeraDeaths) + 
  annotation_map_tile(type = "osm", zoomin = 0) + 
  geom_sf(aes(size = Count), alpha = 0.7)
```

```{r}
library(mapproj)
library(maps)
map("world", projection = "mercator", wrap = TRUE)
map("world", projection = "cylequalarea", param = 45, wrap = TRUE)
```

```{r}
map(
  "state", projection = "lambert", 
  parameters = c(lat0 = 20, lat1 = 50), wrap = TRUE
)
map(
  "state", projection = "albers", 
  parameters = c(lat0 = 20, lat1 = 50), wrap = TRUE
)
```

```{r}
st_crs(CholeraDeaths )
```

There are many CRSs, but a few are most common. A set of EPSG ([*European Petroleum Survey Group*](https://en.wikipedia.org/w/index.php?search=European%20Petroleum%20Survey%20Group)) codes provides a shorthand for the full descriptions (like the one shown above). The most commonly-used are:

[EPSG:4326](https://epsg.io/4326) - Also known as [WGS84](https://en.wikipedia.org/wiki/World_Geodetic_System#WGS84), this is the standard for GPS systems and Google Earth.

[EPSG:3857](https://epsg.io/3857) - A Mercator projection used in maps tiles^[33](https://mdsr-book.github.io/mdsr2e/ch-spatial.html#fn33)^ by Google Maps, Open Street Maps, etc.

[EPSG:27700](https://epsg.io/27700) - Also known as OSGB 1936, or the British National Grid: United Kingdom Ordnance Survey. It is commonly used in Britain.

```{r}
cholera_4326 <- CholeraDeaths_Book %>% 
  st_transform(4326)
```

```{r}
st_bbox(cholera_4326
        )
st_bbox(CholeraDeaths_Book)
```

```{r}
ggplot(cholera_4326) + 
  annotation_map_tile(type = "osm", zoomin = 0) + 
  geom_sf(aes(size = Count), alpha = 0.7)
```

```{r}
cholera_latlong <- CholeraDeaths %>%
  st_set_crs(27700) %>%
  st_transform(4326)
snow <- ggplot(cholera_latlong) + 
  annotation_map_tile(type = "osm", zoomin = 0) + 
  geom_sf(aes(size = Count))

pumps <- st_read(dsn,layer = "Pumps")

pumps_latlong <- pumps %>% 
  st_set_crs(27700) %>%
  st_transform(4326)
snow +
  geom_sf(data = pumps_latlong, size = 3, color = "red")
```

Exercise 1

