---
title: "Schwab's Baby Name Solution"
editor: visual
---


```{r}
library(tidyverse)
library(babynames)
```

## 1. Wrangle your name out of babynames

You will want to `filter()` your name. I would make a vector of names c("Nic","Nicholas", "Nick")

-   In what year was your name the most popular?

-   Make a graph and facet by sex.

```{r}
babynames_Nic <- babynames |>
  filter(name %in% c("Nic","Nicholas","Nick")) |> # I've filtered my name out
  group_by(year)|>
  summarise(prop_total =sum(prop), total = sum(n))  # I've calculated the total of all the Nic's born in a year I'm saving prop because I need it for part 2
  
babynames_Nic
```
## 2. Create a line graph

Show the proportion of children with the same names as you over time.


```{r}
babynames_Nic |>
  ggplot()+
  geom_line(aes(x = year, y= prop_total)) +
  labs(title ="Nicholas born since 1880",
       subtitle =" Data includes Nic, Nick and Nicholas",
       y= "proportion of babies named Nicholas")+
  theme_classic()

```

# Part 2. 5 most popular names in the US since 1880

Do some wrangling to tell me the five most popular names separated by sex.

```{r}
babynames |>
  filter(sex == "M") |>
  group_by(name) |>
  summarise(total = sum(n)) |>
  slice_max(total, n=5)  

babynames |>
  filter(sex =="F") |>
  group_by(name) |>
  summarise(total = sum(n)) |>
  arrange(desc(total)) |>
  slice_head(n=5)
```

# Part 3. Do presidents' names influence babies' names?

Wrangle the data to show how president's first names effect newborn babies' names. [Former Presidents](https://www.whitehouse.gov/about-the-white-house/presidents/)

-   Choose three recent presidents (Donald Trump and Joe Biden might be too recent).

-   Make one graph that shows all three president's names. Use `geom_vline()` to show the year they entered office.


```{r}
babynames |>
  filter(name %in% c("William","George","Barack")) |>
  filter(year > 1990) |>  # I really only care about the years after they were elected. 
  group_by(year, name) |>
  summarize(prop_total = sum(prop), total = sum(n)) |>
  ggplot() +
  geom_line(aes( x= year, y= prop_total, color = name)) +
  geom_vline(xintercept = 2009, color="Red")+
  geom_vline(xintercept = 1993, color="Green")+
  geom_vline(xintercept = 2001, color="Blue")+
  labs(color = "President",
         y = "proportion of new borns")
```

# Part 4: Most and least.

What is the year in which the most babies were born?

What is the year the least babies were born?


```{r}
babynames_total <- babynames |>
  group_by(year)|>
  summarise(total = sum(n))
  
babynames_total |>
  slice_min(total)

babynames_total |>
  slice_max(total)
```