---
title: "2019_2021 Data"
author: "Schwab"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: pdf_document
---

```{r setup, include=FALSE}

#Below I've set it so all your code will output when you knit, because I want to be able to see it. 
knitr::opts_chunk$set(echo = TRUE)

#Here are our libraries
library(tidyverse)
library(readxl)
library(RColorBrewer)

#This is how you import a spreadsheet. the excel file needs to be in the same folder as this .Rmd file
HCC_DATA_19_21 <- read_excel("HCC-DATA-19-21.xlsx")

#Below I am making my excel spreadsheet into a dataframe and saving it to R's environment
HCC_DATA_19_21 <- as.data.frame(HCC_DATA_19_21)

# This is just to make rounding bearable. 
options(scipen=1)

#Making DEGSEEK either yes or no. 
HCC_DATA_19_21$DEGSEEK <- recode_factor(HCC_DATA_19_21$DEGSEEK,"0"="No","1"="Yes")

#Below I am making TERM easier to read
HCC_DATA_19_21$TERM <- recode_factor(HCC_DATA_19_21$TERM,"201930"="2019","202030"="2020","202130"="2021")

#Below I rename RACE so they are easier to read and order. 
HCC_DATA_19_21$RACE <- recode_factor(HCC_DATA_19_21$RACE, 
                               "a_WHITE" = "White",
                               "b_HISPANIC" = "Hispanic",
                               "c_BLACK/NON-HISPANIC" = "Black/Non-Hispanic",
                               "d_ASIAN" = "Asian",
                               "e_AMERICAN IND/ALASKAN"="American Indian/Alaska Native",
                               "f_NATIVE HAWAIIAN" = "Native Hawaiian",
                               "g_CAPE VERDEAN" = "Cape Verdean",
                               "h_TWO OR MORE RACES" = "Two or More Races",
                               "i_UNKNOWN" = "Unknown",
                               "j_NON-RESIDENT" = "Non-Resident"
                               )

#I am tempted to recode GENDER_DESC, but I've left it for you if you want to try it. 

```

```{r look, echo=FALSE, eval=FALSE}
head(HCC_DATA_19_21 )
```

Question 1: 
What is the gender breakdown of degree seeking students? Hint: The count() function works well here.

```{r question_1}

```

Question 2: 
What are the top five towns HCC students are came from in 2019? What percentage of the student body come from the largest town in 2019? 

```{r question_2}

```

Question 3: 
Take the top five towns in 2021 and make a graphical representation that can help to answer the question: How did the Covid-19 pandemic effect the enrollment of HCC students by Race for the top 5 towns from 2019-2021? For bonus use color and add a theme. Also this line of code will rotate your town names, add it to your geom_* : theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

```{r question_3}

```

Question 4: Consider the Data: Make a question and answer it with a graphical display. Add on the labs(),xlab(),ylab() layers to add relevant labels and tiles to your plot. Your plot should be easy to understand.  

```{r question_4}

```


